---
title: "Final Project"
author: '2046727'
date: "1/2/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse) 
library(plyr)
library(dplyr)
library(gridExtra) #for plotting
library(ggrepel) #For plotting
library(naniar) # for missing values
```

Introduction

Over the past few decades consoles for playing video games have become commonplace in houses across the country. Software and hardware advancements of the personal computer (PC) have led to many different machines being developed. Every year, the video game industry churns out hundreds of titles and sells hundreds of millions of units around the globe. Many consoles have come and gone, but a few modern machines have carved their niche in society and continue to evolve and thrive. Xbox (developed by Microsoft), Playstation (developed by Sony), Nintendo, and the personal computer are examples of the platforms that have the biggest hit games from a variety of publishers.
In this report, the following questions of interest are addressed:

• Is there any evidence of differences between sales in different regions of the world?

• Are there any associations when comparing the genre of a game and the platform on which it is released?

• Are there any associations in sales or titles when we consider more current popular platforms?

• Do critic ratings correlate with sales? Do critics tend to favor some publishers/genres/platforms over others?

```{r cars}
summary(cars)
```

Data exploration and cleaning

```{r pressure, echo=FALSE}
##Summary of data
str(games)
summary(games)

##table showing rating
table(games$Rating)

#Rename the abbreviated ratings to their full name
games$Rating[games$Rating=="RP"]="RatingPending"
games$Rating[games$Rating=="EC"]="EarlyChildhood"
games$Rating[games$Rating=="E"]="Everyone"
games$Rating[games$Rating=="E10+"]="Everyone10+"
games$Rating[games$Rating=="T"]="Teen"
games$Rating[games$Rating=="M"]="Mature"
games$Rating[games$Rating=="AO"]="AdultOnly"
games$Rating[games$Rating=="K-A"]="KidsToAdults"

# table with year of release for each game 
table(games$Year_of_Release)

# check for missing values 
gg_miss_var(games)
#There seems to be many missing values and from the descrption of the data set these m
#For the missing values, I think I will leave them as is. Replacing or removing that #many observations might not give the most meaningful results 

#check for duplication 
duplicated(games) %>%
    sum()
```

What are the top selling games by region and their platform & genre?
```{r}
# For this plot the Top Games are considered the ones with more than 20 million sales
ggplot(games[games$Global_Sales>20,], aes(x=reorder(Name,Global_Sales), y=Global_Sales, 
                                            fill=Platform)) +
    geom_bar(stat="identity")+
    theme(legend.position="top")+
    labs(title="Top Selling Games Globally", x= "Game and Genre", 
         y= "Global Sales in Millions")+
    geom_text(aes(label=Genre), hjust=1.1,color="white", size=3)+ 
    coord_flip()
```
what are the top 8 games per region?
```{r}
p1=ggplot(games[games$JP_Sales>5.33,], aes(x=reorder(Name,JP_Sales), y=JP_Sales, 
                fill=Platform))+
     geom_bar(stat="identity")+
    labs( x= "Game and Genre", y= "Japan Sales in Millions")+
    geom_text(aes(label=Genre), hjust=1.1, color="white", size=3)+
    coord_flip()

p2=ggplot(games[games$NA_Sales>14,], aes(x=reorder(Name,NA_Sales), y=NA_Sales, 
                                            fill=Platform)) +
    geom_bar(stat="identity")+
    labs(x= "Game and Genre", y= "North America Sales in Millions")+
    geom_text(aes(label=Genre), hjust=1.1,color="white", size=3)+
    coord_flip()

p3=ggplot(games[games$EU_Sales>9.0,], aes(x=reorder(Name,EU_Sales), y=EU_Sales, 
                                            fill=Platform)) +
    geom_bar(stat="identity")+
    labs(x= "Game and Genre", y= "Europe Sales in Millions")+
    geom_text(aes(label=Genre), hjust=1.1,color="white", size=3)+
    coord_flip()

p4=ggplot(games[games$Other_Sales>2.85,], aes(x=reorder(Name,Other_Sales), y=Other_Sales, 
                                            fill=Platform)) +
    geom_bar(stat="identity")+
    labs(x= "Game and Genre", y= "Sales of the Other Regions in Millions")+
    geom_text(aes(label=Genre), hjust=1.1,color="white", size=3)+
    coord_flip()

grid.arrange(p1,p2,p3,p4, nrow=4)
```
What is the total number of games released every year?¶
```{r}
ggplot(games[games$Year_of_Release!="N/A",], aes(x=Year_of_Release, fill=..count..)) +
    geom_bar()+
    scale_color_gradient(low="purple", high="orange")+
    scale_fill_gradient(low="purple", high="orange")+
    labs(title="Number of Games Released every Year", x= "Game", 
         y= "Total Number of Games")+
    geom_text(stat='count',aes(label=..count..), hjust=-0.1,color="darkblue", size=2.5)+
    coord_flip()
```
How has the sales of games changed throughout the years?
```{r}
yearly=games %>%
    group_by(Year_of_Release) %>%
    dplyr::summarize(numrelease=n(),Japan=sum(JP_Sales), NorthAmerica=sum(NA_Sales),
                     Europe=sum(EU_Sales),Others=sum(Other_Sales), 
                    Worldwide=sum(Global_Sales)) 

yearly$Year_of_Release = as.numeric((yearly$Year_of_Release))
yearly=yearly[!(is.na(yearly$Year_of_Release)),]

#There are only 3 observations in the year 2017 so I will remove them
yearly=subset(yearly, yearly$Year_of_Release!=2017)


years = yearly %>% 
  gather(Region, Value, -c(Year_of_Release, numrelease)) %>% 
  arrange(Region)

ggplot(years, aes(x=Year_of_Release, y=Value, color=Region))+
    geom_point()+
    geom_line()+
    theme(legend.position="top")+
    theme(axis.text.x = element_text(angle=60, hjust=1)) +
    scale_x_continuous("Year",labels = as.character(years$Year_of_Release), 
                       breaks=years$Year_of_Release)
```
What are the highest global selling games per decade?
To answer this, I will first need to create a new variable called Decade which will have 4 values, and assign each game based on their year of release to their appropriate decade. Then I'll display the results in a circular bar chart
```{r}
topglobal= games %>%
    arrange(desc(Global_Sales)) %>%
    group_by(Decade) %>%
    slice(1:5)%>%
    filter(!is.na(Decade)) #removed data with missing decade (since year of release was N/A)
topglobal$NamePlat=paste(topglobal$Name, topglobal$Platform, sep=" -")
topglobal=topglobal %>%
    arrange(match(Decade, c("2010s","2000s", "90s","80s")), 
            desc(Global_Sales))
topglobal$id=as.numeric(seq(1,20)) #id for the display of each value on the circular barchart
topglobal$Decade=factor(topglobal$Decade,
                        levels=c("2010s","2000s", "90s","80s"))

games$Decade[games$Year_of_Release>=1980 & games$Year_of_Release<=1989]="80s"
games$Decade[games$Year_of_Release>=1990 & games$Year_of_Release<=1999]="90s"
games$Decade[games$Year_of_Release>=2000 & games$Year_of_Release<=2009]="2000s"
games$Decade[games$Year_of_Release>=2010 & games$Year_of_Release<=2019]="2010s"
```
Which publishers have the highest game sales?
```{r}
pub = games %>%
    group_by(Publisher) %>%
    dplyr::summarize(total=n(), salestotal=round(sum(Global_Sales),1), 
                    salesjp=round(sum(JP_Sales),1),
                    salesna=round(sum(NA_Sales),1),
                    saleseu=round(sum(EU_Sales),1),
                    other=round(sum(Other_Sales),1)) 

#In this plot the highest game sales are considered the ones with more than 100 million sales
ggplot(pub[pub$salestotal>100,], aes(x=reorder(Publisher,salestotal), y=salestotal)) +
    geom_point(aes(color=salestotal), shape=19, size=3)+
    scale_color_gradient(low="skyblue", high="purple")+
    geom_segment(aes(x=Publisher, xend=Publisher, y=0,yend=salestotal, color=salestotal), 
                 size=1) + 
    theme(legend.position="top")+
    labs(title="Publishers with Top Selling Games Worldwide", x= "Publisher", 
         y= "Total Global Sales in Millions")+
    theme_bw()+
    geom_text_repel(aes(label=salestotal),size=3,data=pub[pub$salestotal>100,],
                    color="black")+ 
    coord_flip()




pt1 =ggplot(pub[pub$salesjp>50,], aes(x=reorder(Publisher,salesjp), y=salesjp)) +
    geom_point(aes(color=salesjp), shape=19, size=3)+
    geom_segment(aes(x=Publisher, xend=Publisher,y=0,yend=salesjp,color=salesjp), size=1)+ 
    labs( x= "Publisher", y= "Total Japan Sales in Millions")+
    theme_bw()+
    scale_color_gradient(low="skyblue", high="purple")+
    coord_flip()

pt2 = ggplot(pub[pub$salesna>110,], aes(x=reorder(Publisher,salesna), y=salesna)) +
    geom_point(aes(color=salesna), shape=19, size=3)+
    geom_segment(aes(x=Publisher, xend=Publisher,y=0,yend=salesna,color=salesna), size=1)+ 
    labs( x= "Publisher", y= "Total North America Sales in Millions")+
    theme_bw()+
    scale_color_gradient(low="skyblue", high="purple")+
    coord_flip()

pt3 = ggplot(pub[pub$saleseu>80,], aes(x=reorder(Publisher,saleseu), y=saleseu)) +
    geom_point(aes(color=saleseu), shape=19, size=3)+
    geom_segment(aes(x=Publisher, xend=Publisher,y=0,yend=saleseu,color=saleseu), size=1)+ 
    labs( x= "Publisher", y= "Total Europe Sales in Millions")+
    theme_bw()+
    scale_color_gradient(low="skyblue", high="purple")+
    coord_flip()

pt4 =  ggplot(pub[pub$other>40,], aes(x=reorder(Publisher,other), y=other)) +
    geom_point(aes(color=other),shape=19, size=3)+
    geom_segment(aes(x=Publisher, xend=Publisher,y=0,yend=other,color=other),
                 size=1)+ 
    labs( x= "Publisher",y= "Total Sales in Other Regions in Millions")+
    theme_bw()+
    scale_color_gradient(low="skyblue", high="purple")+
    coord_flip()

grid.arrange(pt1,pt2,pt3,pt4, nrow=4)
```
What are the most famous gaming genres released by the top publishers?
Here what I want to find is what are the gaming genres that the top selling publishers release. From the plot above,I will use the top 6 selling publishers worldwide which are Nintendo, Electronic Arts, Activision, Sony Computer Entertainment , Ubisoft and Take-Two Interactive.
```{r}
gpub= games %>%
    group_by (Publisher, Genre) %>%
    dplyr:: summarize(total=n(), sumsales=round(sum(Global_Sales),1)) %>%
    filter(Publisher %in% c("Nintendo", "Electronic Arts", "Activision", 
                      "Sony Computer Entertainment","Ubisoft", "Take-Two Interactive")) 

gpub$Publisher = factor(gpub$Publisher, 
                               levels=c("Nintendo", "Electronic Arts", "Activision", 
                      "Sony Computer Entertainment","Ubisoft", "Take-Two Interactive"))

ggplot( gpub, aes(x=Genre, y=total, fill=Genre))+
    geom_bar(stat="identity") +
    facet_wrap(~Publisher, ncol=2, scale="free") +
    theme(legend.position= "none")+
    labs(title="Total Number of Games Released per Genre for the Top 6 Publishers",
         x= "Genre", 
         y= "Total Number of Games Released")+
    theme(axis.text.x=element_text(angle=60, hjust=1, size=8))+
    geom_text(aes(label=total), vjust=0.5,color="black", size=2.25)
```
